---
title: "Data analysis - Rodger Farr"
author: "Maxwel C Oliveira"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, echo=FALSE, cache=FALSE}
library(knitr)
library(rmdformats)

## Global options
options(max.print = .Machine$integer.max)
opts_chunk$set(echo=TRUE,
               cache=TRUE,
               fig.cap='center',
               prompt=FALSE,
               tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=300)
```

# Load packages

```{r message=FALSE, ec}
library(lme4)
library(lmerTest)
library(tidyverse)
library(emmeans)
library(kableExtra)
```


# Load data

```{r message=FALSE}
data <- read_csv("iwm.csv") %>% 
  mutate(
    year = as.factor(year),
    trt = as.factor(trt)
  ) %>% 
  mutate_if(is_character, as_factor)
```


# Check your data

```{r}
glimpse(data) 
```

```{r}
skimr::skim(data)
```


# Assumptions


## Homogeneity of variances


```{r}
bartlett.test(yield ~ trt, data)
```



## Normality

Testing and understanding normality is tricky. There are a lot of misunderstanding in the literature. I believe you should not worry much about it.

```{r}
data %>% 
  ggplot(aes(x=yield)) +
  geom_histogram(bins = 10)
```

```{r}
library(ggridges)
data %>% 
  ggplot(
    aes(y=as.factor(trt), x=yield, fill=as.factor(trt))) +
      geom_density_ridges(scale=2, show.legend = FALSE)
```

# Summarising your data

```{r}
data %>%
  group_by(hand.weed, plow, cover.crop, herb, year) %>% 
  summarise(yield_mean = mean(yield), yield_sd = sd(yield), 
            yield_min = min(yield), yield_max = max(yield)) %>%
  kbl() %>%
  kable_classic_2(full_width = F)
```



# Raw data visualization

- Because you have several factors, I have used trts to look at your raw data visualization.

- I am using trt and looking over the years. It seems that year play an important role in your study. I meant that there's no much difference among trts but you see a difference between years.



```{r}
data %>% 
  ggplot(aes(x=as.factor(trt), y=yield, fill=year)) +
  geom_bar(position = "dodge", stat="identity")
```


# Yield analysis

## Model

Split plot is a factorial design. I have never needed to adjust a a model to split plot, in your case you have so many. This is the approach I would recommend. I am not an expert in stats, the model that I have built it is the one I would go with your data. Maybe the random effects is wrong but it might be very close.

```{r message=FALSE}
fit <- lmer(yield ~ hand.weed * plow * cover.crop * herb * year  + (1|hand.weed:plow:cover.crop:herb) + (1|year:block), data, REML= FALSE)
```


## Anova

```{r max.print=100000}
anova(fit)
```

## Least square means

### Three-way interaction

- hand.weed:plow:year

```{r}
lsmeans_1 <- emmeans(fit, ~ hand.weed * plow | year, cont="pairwise", adjust="none", alpha=0.05)
lsmeans_1 # note that | make the hand.weed * plow comparisons within year (2 sets of comparisons) - You can play with | or *
```

### Compact letter display


```{r}
CLD(lsmeans_1$emmeans, alpha=0.05, Letters=letters, adjust="none", reversed = TRUE) %>%
  kbl() %>%
  kable_classic_2(full_width = F)
```



## Two-way interaction

- herb:year 

```{r warning=FALSE}
lsmeans_2 <- emmeans(fit, ~ herb * year, cont="pairwise", adjust="none", alpha=0.05)
lsmeans_2
```

### Compact letter display


```{r}
CLD(lsmeans_2$emmeans, alpha=0.05, Letters=letters, adjust="none", reversed = TRUE) %>%
  kbl() %>%
  kable_classic_2(full_width = F)
```


## Finish the other interactions